---
title: "fars_markdown"
author: "A M Mahedi Hasan"
date: "31 August 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fars Functions:

## fars_read function

The "fars_read" function reads a CSV file, given a file name, without showing a progress line.
Throws an error line if the file doesn't exist. Name of the file is the only input parameter and it returns a *tibble*. 

```{r fars_read}
fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}

```
Example: `fars_read("../data/accident_2013.csv.bz2")`

```{r echo=FALSE}
fars_read("C:/Workstation/coursera/Mastering_R/final_assignment_building_R_package/package/farsFinal/data/accident_2013.csv.bz2")
```


## make_filename function

The "make_filename" function creats a file name based on the year that is provided as a parameter. It returns a string, which is the file name with *bz2* zipped extension. 

```{r make_filename}
make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("C:/Workstation/coursera/Mastering_R/final_assignment_building_R_package/package/farsFinal/data/accident_%d.csv.bz2", year)
}
```
Example: `make_filename(2013)`
```{r echo=FALSE}
make_filename(2013)
```


## fars_read_years function

The "fars_read_years" function reads a range of valid years, calls the make_filename function and reads data in batch mode. It returns a list of tibble objects given the range of valid years. If a valid year is not given, the function throws an error message.


```{r}
require(magrittr)
fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

```
Example: `fars_read_years(2013:2014)`

```{r echo=FALSE}
fars_read_years(2013:2014)
```


## fars_summarize_years function

The "fars_summarize_years" function counts the occurance by years and months, calls the "fars_read_years" function to do so. It takes a range of integers (years) as input and returns a wide format tibble object, given the range of valid years, of the count of occurance by years and months. 


```{r}
library(magrittr)
fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by(year, MONTH) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}
```

Example: `fars_summarize_years(2013:2014)`

```{r echo=FALSE}
fars_summarize_years(2013:2015)

```


## fars_map_state function

The "fars_map_state" function draws a map for a given state number in the data and a valid year. It returns a *map*. It throws an error "invalid STATE number:" if an invalid state number is given. If no data is available, it throws a message "no accidents to plot". 

```{r}
fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)
  
  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}
```

Examples `fars_map_state(1, 2013)`
```{r echo=FALSE}
fars_map_state(1, 2013)

```




